A Study of Institutional Spending on Open Access Publication Fees in Germany

### 2016-06-10 article.Rmd
fix superscript in scoap3

This study is limited in some respects. First, we cannot assess whether publishers and journals granted publication fee discounts. The Open APC initiative uses a minimal data scheme to encourage self-reporting, and it therefore does not track this type of information. However, large price variations suggest that different pricing levels and pricing schemes are in place, as previously observed [@Solomon_2012; @Pinfield_2015]. Adding to this complexity, it is likely that some institutions paid only part of the publication fee. Consider, for instance, the journal Nature Communication: Charges reported to the Open APC initiative ranged between € 2000, the DFG price cap, and € 4,403 for a journal that can be categorised as a pricy, high-impact journal according to @Solomon_2012. Although making such payments from several budgets is a proposed strategy to sustain publication funds at German universities [@Fournier_2013], these pro-rata payments were not made transparent in the Open APC data, leading to a possibly flawed representation of publication fee spending in Germany. In another case, one university included its contributions to the SCOAP^3 consortia and presumably divided the sum by the number of articles published by their authors in SCOAP^3-covered journals.[^26] This approach is arbitrary, because averages for an institution can be determined only after the end of a full 3-year funding cycle. Other factors affecting price variations are exchange rates and different tax rates for some organisations in Germany. For instance, the Max Planck Society has a limited input tax reduction. The refund of input value-added tax for publication fees is 20 %. To increase the transparency of publication fee spending, @Pinfield_2015 suggested disclosing tax rates and payment currencies in future cost data-sets. In addition, the Open APC data-set does not track funding sources; thus, we could not determine which funders co-financed publication fees.

@@

[^26]: SCOAP^3, the Sponsoring Consortium for Open Access Publishing in Particle Physics, is a unique approach to convert former subscription journals in high-energy physics to open access journals under a CC BY license. The consortiums, led by CERN, pays publishers centrally, based on previously agreed APCs and an overall price cap, and retrieves its funds from organisations and countries based on their share in the articles published in the covered journals. German universities participate through an initiative led by the German National Library of Science and Technology that received additional funding from DFG.

>>>>

This study is limited in some respects. First, we cannot assess whether publishers and journals granted publication fee discounts. The Open APC initiative uses a minimal data scheme to encourage self-reporting, and it therefore does not track this type of information. However, large price variations suggest that different pricing levels and pricing schemes are in place, as previously observed [@Solomon_2012; @Pinfield_2015]. Adding to this complexity, it is likely that some institutions paid only part of the publication fee. Consider, for instance, the journal Nature Communication: Charges reported to the Open APC initiative ranged between € 2000, the DFG price cap, and € 4,403 for a journal that can be categorised as a pricy, high-impact journal according to @Solomon_2012. Although making such payments from several budgets is a proposed strategy to sustain publication funds at German universities [@Fournier_2013], these pro-rata payments were not made transparent in the Open APC data, leading to a possibly flawed representation of publication fee spending in Germany. In another case, one university included its contributions to the SCOAP^3^ consortia and presumably divided the sum by the number of articles published by their authors in SCOAP^3^-covered journals.[^26] This approach is arbitrary, because averages for an institution can be determined only after the end of a full 3-year funding cycle. Other factors affecting price variations are exchange rates and different tax rates for some organisations in Germany. For instance, the Max Planck Society has a limited input tax reduction. The refund of input value-added tax for publication fees is 20 %. To increase the transparency of publication fee spending, @Pinfield_2015 suggested disclosing tax rates and payment currencies in future cost data-sets. In addition, the Open APC data-set does not track funding sources; thus, we could not determine which funders co-financed publication fees.

@@

[^26]: SCOAP^3^, the Sponsoring Consortium for Open Access Publishing in Particle Physics, is a unique approach to convert former subscription journals in high-energy physics to open access journals under a CC BY license. The consortiums, led by CERN, pays publishers centrally, based on previously agreed APCs and an overall price cap, and retrieves its funds from organisations and countries based on their share in the articles published in the covered journals. German universities participate through an initiative led by the German National Library of Science and Technology that received additional funding from DFG.
